Semistochastic importance sampling of second-quantized operators in determinant space: application to multireference perturbation theory

### 2021-07-01 notes/paper.tex
Rewrite alias sampling module, old SHCI working correctly now

It should be noted that a crude version of semistochastic importance sampling has been used before in the context of FCIQMC. However, in that case, the deterministic component was pre-computed, and the stochastic component was not disjoint from it: there was some probability of sampling an excitation that already existed in the deterministic component and therefore had to be discarded.

In this paper, we describe a unified approach to semistochastic importance sampling, in which the deterministic component is computed efficiently on the fly, and the stochastic component efficiently samples only the remaining terms left out of the deterministic component. We focus on the nonrelativistic quantum chemistry Hamiltonian, but note that the techniques could easily be generalized to alternative Hamiltonians or other second-quantized operators such as cluster operators.

>>>>

Whether deterministic, stochastic, or semistochastic, we define an \emph{efficient} algorithm for applying a second-quantized operator to a Slater determinant as one which does not spend too much time on components of the operator with small coefficients.
\begin{itemize}
\item In the case of deterministic methods, an efficient method would loop over only the large-magnitude terms in the operator, to avoid wasting time on the vast number of small-magnitude terms. HCI, introduced by one of us, made use of this approach to efficiently select new determinants in Selected CI.
\item In the case of stochastic methods (or the stochastic component of semistochastic methods), rather than skipping the small-magnitude terms entirely, an efficient approach simply means not over-sampling small-magnitude terms. That is, the small-magnitude terms are importance-sampled, with probability a function of their magnitude.~\footnote{While a semistochastic extension of HCI (called SHCI) has already been introduced by one of us, it did not have this efficient property; small-magnitude contributions were vastly over-sampled, since the sampling was done according to a distribution that did not take the matrix element magnitude into account.}
\end{itemize}

In this paper, we describe a unified approach to semistochastic importance sampling. Our approach extends the deterministic HCI algorithm to semistochastic methods, in a way that (unlike SHCI) is consistent with the underlying philosophy of HCI: we importance-sample the small-magnitude terms that were not already treated deterministically, thus avoiding wasteful over-sampling of these terms entirely. ~\footnote{It should be noted that a crude version of semistochastic importance sampling has been used before in the context of FCIQMC. However, in that case, the deterministic component was pre-computed, and the stochastic component was not disjoint from it: there was some probability of sampling an excitation that already existed in the deterministic component and therefore had to be discarded.}
%We compute the deterministic component efficiently on the fly, and treat the stochastic component by efficiently sampling only the remaining terms left out of the deterministic component.
We focus on the nonrelativistic quantum chemistry Hamiltonian, but note that the techniques could easily be generalized to alternative Hamiltonians or other second-quantized operators such as cluster operators.
